### 2014-05-07 ApTranscriptome_TR.Rmd
More updates to analysis

## Thermally-responsive transcripts

@@

png(file = paste(resultsdir, "responsive_transcripts_mosaicplot.png", sep = ""))
mosaicplot(tt2, shade = TRUE, xlab = "ApVT", ylab = "AcNC", main = "Mosaic plot of responsive transcripts")
dev.off()

@@

Table 4 shows the number of transcripts that fall into each expression type for each each colony. The totals for each colony include the `r nrow(temperature.lms)` transcripts that have consistent temperature responses between the two colonies.

@@

Selection may have acted in response to thermal stress such that some genes are constitutively activated in one colony while plastically-expressed in the other. I evaluate this hypothesis by comparing expression levels at the at optimum (19.5C) temperature between the two colonies for genes in that are either in the 'High' or 'Low' expression group in the other colony. Specifically, I expect that genes upregulated at high temperatures in *A22* are more highly expressed at 19.5C in *Ar*? Conversely, I expect genes that are upregulated at low temps in *Ar* to have greater constitutive expression in *A22*?

@@

Yes - transcripts that are up-regulated at high temperatures in *A22* are more highly-expressed at 19.25C in *Ar*. Note that A22 had the larger library size so if this was due to TPM not correctly accounting for differences in reads between samples, we would expect to see a positive instead of negative value here.

@@

# t-test
t4 <- t.test(varshape.out$s.slope, alternative = "two.sided")

# curvature = curvature1 - curvature2
varshape.out$curvature <- varshape.out$A22.curve - varshape.out$Ar.curve
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.curvature <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.curvature"] <- (varshape.out[i, "curvature"] / mean(c(varshape.out[i, "A22.curve"], varshape.out[i, "Ar.curve"])))
}
# change 'Inf' values to 0
varshape.out[which(varshape.out$Ar.curve == 0 & varshape.out$A22.curve == 0), "s.curvature"] <- 0
# t-test
tt.test(varshape.out$s.curvature, alternative = "two.sided")

@@

# t-test
t.test(varshape.out$s.wiggle, alternative = "two.sided")
```

@@

t.test(varshape.out$s.total, alternative = "two.sided")

@@

## Functional annotation ##

@@

There are also 6 heat shock related genes in the responsive transcripts, out of 130 total.

@@

# table results
Ap.BP.ResTable <- GenTable(Ap.BP.GOdata, parentchild = Ap.BP.resultParentChild, topNodes = 113)
dim(Ap.BP.ResTable)
#pandoc.table(Ap.BP.ResTable)

>>>>

### Grouping of thermally-responsive transcripts

@@

png(file = paste(resultsdir, "responsive_transcripts_mosaicplot.png", sep = ""))
mosaicplot(tt2, xlab = "ApVT", ylab = "AcNC", main = "Mosaic plot of responsive transcripts")
dev.off()

@@

Table 4 shows the number of transcripts that fall into each expression type for each each colony. The totals for each colony include the `r length(temperature.lms)` transcripts that have consistent temperature responses between the two colonies.

@@

Selection may have acted in response to thermal stress such that some genes are constitutively activated in one colony while plastically-expressed in the other. I evaluate this hypothesis by comparing expression levels at the at optimum (19.5C) temperature between the two colonies for genes in that are either in the 'High' or 'Low' expression group in the other colony. Specifically, I expect that genes upregulated at high temperatures in *A22* are more highly expressed at 19.5C in *Ar*?

@@

Consistent with expectation, there is greater expression at 19.5C for *Ar* than *A22* transcripts for the set of transcripts that are transcripts that are up-regulated at high temperatures in *A22*. Note that A22 had the larger library size so if this was due to TPM not correctly accounting for differences in reads between samples, we would expect to see a positive instead of negative value here.

@@

# t-test
(t4 <- t.test(varshape.out$s.slope, alternative = "two.sided"))

# curvature = curvature1 - curvature2
varshape.out$curvature <- varshape.out$A22.curve - varshape.out$Ar.curve
# to compare among transcripts that differ in scale of expression, standardize by overall mean for each transcript
varshape.out$s.curvature <- NA
for(i in 1:nrow(varshape.out)) {
  varshape.out[i, "s.curvature"] <- (varshape.out[i, "curvature"] / mean(c(varshape.out[i, "A22.curve"], varshape.out[i, "Ar.curve"])))
}
# change 'Inf' values to 0
varshape.out[which(varshape.out$Ar.curve == 0 & varshape.out$A22.curve == 0), "s.curvature"] <- 0
# t-test
(t.test(varshape.out$s.curvature, alternative = "two.sided"))

@@

# t-test
(t5 <- t.test(varshape.out$s.wiggle, alternative = "two.sided"))
```

@@

(t6 <- t.test(varshape.out$s.total, alternative = "two.sided"))

@@

## Functional annotation

@@

There are also 7 heat shock related genes in the responsive transcripts, out of 130 total.

@@

# table results
Ap.BP.ResTable <- GenTable(Ap.BP.GOdata, parentchild = Ap.BP.resultParentChild, topNodes = 137)
dim(Ap.BP.ResTable)
#pandoc.table(Ap.BP.ResTable)
